Reconstructing Propositional Proofs in Type Theory

### 2017-10-31 paper/sections/type-theory.tex
[ paper ] some fixes in type-theory section.

We produce derivation trees inside the judgments using the derivation
rules.  Since type theory can be seen as typed $λ$-calculus with
dependent function types, evaluation of $λ$-terms also called
\emph{normalization} of the proof-terms, are reductions using the
system inference rules. Therefore, the task to verify a proof that
proves a certain theorem is translated to verify if the associated
tree of the proof has the correspondent type of the theorem. We know
this process as \emph{type-checking}. This feature of type theory
allow us to verify a proof generated by an \ATP by reconstructing its
proof-term to type-check the proof. For such a purpose, we use a proof
assistant in type theory like \Agda to delegate this task.  However,
this approach has some weak points discussed in~\cite{Bezem2002}. Some
limitations from the type theory point of view for
proof-reconstruction is described in the following sections.

@@

Algorithms based on general recursion can not be translated
directly in type theory since it is not a guarantee they terminate. For that
reason, we follow the technique described in~\cite{Bertot2004} to avoid
termination problems by modifying the recursive functions to be
\emph{structural recursive}.
% (see \cite{Coquand1992,Abel2002,Bove2005} for more details).

@@

\begin{notation}
  We use \Prop for the type of propositions.  For syntactical
  equivalence, we write the equality as $φ ≡ ψ$ for $φ,~ψ :
  \Prop$. For this equality, we assume the reflexivity, symmetry and
  transitivity properties. The type of natural numbers is called \Nat,
  and it is defined as usual, where $\zero$ and $\suc$ are its data
  constructors. We use names and symbols for the arithmetic operations
  as usual ($+$, $-$, $*$).  We use syntax sugar for $\zero$, $\suc$,
  $\suc~\zero$, $\dots$, with the decimal representation~$0$, $1$,
  $2$, $\dots$ as well.
\end{notation}

Hence, one approach to define a structural recursive function based on
a general recursive function $\rm{f}:~\Prop~\to~\Prop$ is to formulate a new
function $\rm{f}^{*}~:~\Prop~\to~\Nat~\to~\Prop$ where all of its recursive
calls are done on smaller arguments.

@@

We bounded the recursion calls in~\eqref{eq:uh-structured} with a \Nat argument. This natural number represents the complexity measure of $\fuh_{0}$
function.

The \emph{complexity measure} defined in \cite{Agudelo-Agudelo2017}
and presented in \cite{VanDalen1994} as \emph{the rank for a
proposition}, is a function $\Prop \to \Nat$ that assigns a natural
number to a formula like we do with the following $\frank$ function.
The symbols $c_{¬}$ and $c_{\square}$ are non-negative integer
numbers called the complexity of the respective connective.

\begin{equation}
% \label{eq:rank-definition}
\begin{aligned}
&\hspace{.495mm}\frank : \Prop → \Nat\\
&\begin{array}{llll}
\frank~(¬~φ)           &= \frank~φ₁~+~c_{¬}\\
\frank~(φ₁~\square~φ₂) &= \fmax~(\frank~φ₁)~(\frank~φ₂) + c_{\square}\\
\frank~φ               &= 0.
\end{array}
\end{aligned}
\end{equation}

We define the $\fuh_{cm}: \Prop → \Nat$ function as
the complexity measure for the $\fuh_0$ function.

>>>>

We produce derivation trees inside the judgments using the derivation rules.
Since type theory can be seen as typed $λ$-calculus with dependent function
types, evaluation of $λ$-terms also called \emph{normalization} of the proof-
terms, are reductions using the system inference rules. Therefore, the task to
verify a proof that proves a certain theorem is translated to verify if the
associated tree of the proof has the correspondent type of the theorem. We know
this process as \emph{type-checking}. This feature of type theory allow us to
verify a proof generated by an \ATP by reconstructing its proof-term to type-
check the proof. For such a purpose, we use a proof assistant based on type
theory like \Agda to delegate this task.
% However, this approach has some weak points discussed in~\cite{Bezem2002}.
Some limitations from the type theory
point of view for proof-reconstruction is described in the following sections.

@@

Functions defined by a general recursion can not be directly translated in type
theory since it is not a guarantee they terminate. For that reason, we follow
the technique described in~\cite{Bertot2004} to avoid termination problems by
modifying the recursive functions to be \emph{structural recursive}. % (see
\cite{Coquand1992,Abel2002,Bove2005} for more details).

@@

\begin{notation}
  We use \Prop for the type of propositions.  For syntactical
  equivalence, we write the equality as $φ ≡ ψ$, for $φ,~ψ :
  \Prop$. For this equality, we assume the reflexivity, symmetry and
  transitivity properties. The type of natural numbers is called \Nat,
  and it is defined as usual, where $\zero$ and $\suc$ are its data
  constructors. We use names and symbols for the arithmetic operations
  as usual (\eg, $+$, $-$, $*$).  We use syntax sugar for $\zero$, $\suc$,
  $\suc~\zero$, $\dots$, with the decimal representation~$0$, $1$,
  $2$, $\dots$ as well.
\end{notation}

Hence, one approach to define a structural recursive function based on a general
recursive function $\rm{f}:~\abbre{a}~\to~\abbre{b}$ is to formulate a new
function $\rm{f}^{*}~:~\abbre{a}~\to~\Nat~\to~\abbre{b}$ where all of its
recursive calls are done on smaller arguments.

@@

We bounded the recursion calls of the $\fuh_{0}$ function using its
\emph{complexity measure}.

The complexity measure of a function is the number of steps
the function takes to finish. Since the functions defined for formulas
are based on the formula structure, we assign a complexity measure for each
of these cases in the formula structure. We refer the reader to~\cite{VanDalen1994,Agudelo-Agudelo2017} for more details about this kind of function.

In our case, we define the $\fuh_{cm}: \Prop → \Nat$ function as
the complexity measure for the $\fuh_0$ function.
